---
title: "Apply cTSI to Escambia Bay"
author: "J. Hagy"
date: "January 25, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("ggmap")

library(tidyverse)
library(gridExtra)
library(lubridate)
library(caret)
library(sf)
library(maps)
library(ggmap)
library(mapview)
library(leaflet)
library(RColorBrewer)
library(qgam)
select <- dplyr::select

source("functions/coastal_tsi_functions.R")


```

## Read Boston Harbor Data
Data downloaded from: https://www.mwra.com/harbor/html/wq_data.htm

```{r}

includeStations <- c(24,38,106,124,138,139,140,141,142)

bh_station_locations <- read.csv("boston_harbor_sampling_data_locations.csv",
                                 stringsAsFactors = FALSE) %>% 
  rename(Station.ID=STAT_ID,
         Latitude=TARGET_LAT,
         Longitude=TARGET_LON) %>% 
  mutate(Station.ID=as.numeric(Station.ID)) %>% 
  filter(Station.ID %in% includeStations,
         STUDY_ID=="BHWQM") %>% 
  st_as_sf(., coords = c('Longitude', 'Latitude'), crs = 4326) %>% 
  mutate(type="Nutrients")


# note: stations with letters in them aren't needed, so doesn't matter that they can't be coerced 
# to numeric.

```

# Make a map of the Boston Harbor stations

```{r}

pal <- colorFactor(
  palette = brewer.pal(3,"Set2"),
  domain = bh_station_locations$type)

m  <-  leaflet(bh_station_locations, height = 350*1, width = 350*1, options = leafletOptions(zoomControl = FALSE)) %>% 
  addProviderTiles("CartoDB.PositronNoLabels") %>%
  setView(-70.95, 42.3, zoom = 11) %>%
  addCircleMarkers(weight=2,radius=6,opacity=1,
                   label = ~Station.ID, 
                   labelOptions = labelOptions(noHide = TRUE, offset=c(18,18), 
                      textOnly = TRUE, direction = "top", 
                      style = list("color" = "black","font-weight" = "bold",
                                   "font-size" = "9px"))) %>% 
  addScaleBar("bottomleft") %>%
  addMiniMap(
    tiles = providers$CartoDB.PositronNoLabels,
    position = "topright",
    centerFixed=c(40,-72),
    width = 80, height = 110, zoomLevelOffset = -7,
    toggleDisplay = FALSE)
m
#m

mapshot(m, file = "Stations.jpg", vheight=350, vwidth=350, zoom=3, remove_controls = FALSE)

# Note:  needed to install PhantomJS to make this work, as below
#PhantomJS not found. You can install it with webshot::install_phantomjs(). If it is installed, #please make sure the phantomjs executable can be found via the PATH variable.
#> webshot::install_phantomjs()
#trying URL 'https://github.com/wch/webshot/releases/download/v0.3.1/phantomjs-2.1.1-windows.zip#'
#Content type 'application/octet-stream' length 18193653 bytes (17.4 MB)
#downloaded 17.4 MB
#phantomjs has been installed to C:\Users\jhagy\AppData\Roaming\PhantomJS

#bh_shorelines <- st_read("C:/Users/jhagy/OneDrive - Environmental Protection Agency #(EPA)/jim/GIS/bostonharbor/High Water Shorelines (1800s-2009)/

```
# Make a map using ggmap
```{r}

bh_station_locations <- read.csv("boston_harbor_sampling_data_locations.csv",
                                 stringsAsFactors = FALSE) %>% 
  rename(Station.ID=STAT_ID,
         Latitude=TARGET_LAT,
         Longitude=TARGET_LON) %>% 
  mutate(Station.ID=as.numeric(Station.ID)) %>% 
  filter(Station.ID %in% includeStations,
         STUDY_ID=="BHWQM")

bh_bbox <- make_bbox(lat = Latitude, lon=Longitude,data=bh_station_locations)

names(bh_bbox) <- c("left","bottom","right","top")

bh_map <- get_map(location = bh_bbox, maptype = "watercolor", source = "stamen",zoom=12)
ggmap(bh_map) +
  geom_point(data=bh_station_locations,mapping = aes(x=Longitude,y=Latitude),color="red",shape=18,size=4) +
  
```
# Read the Boston Harbor Nutrients data

```{r}
bh_nutrients <- read.csv("C:/Users/jhagy/OneDrive - Environmental Protection Agency (EPA)/jim/Kreakie Collaborations/Boston Harbor/bh_nutrients.csv",skip=4) %>% 
  filter(Surface.or.Bottom=="S") %>% 
  mutate(TimeStamp=as.POSIXct(Date.time..EASTERN.STANDARD.TIME.,format="%m/%d/%y %H:%M",tz="Etc/GMT-5"))
         
# Select columns by number
keepCols <- c(2,3,4,5,33,8,9,10,11,12,13,14,15,16,19,20,21,22,23,24,25,26,29,30)
bh_nutrients <- bh_nutrients[,keepCols]
# Rename variables - all concentrations are uM, depth in meters
names(bh_nutrients) <- c("Region,","Subregion","DEP.segment","Station.ID","TimeStamp","Sample.Depth","NH4","NH4.Flag","NOx","NOx.Flag","TDN","TDN.Flag","PN","PN.Flag","PO4","PO4.Flag","TDP",
  "TDP.Flag","PP","PP.Flag","TP","TP.Flag","Chla","Chla.Flag")

# Calculate DIN,TN and TP
bh_nutrients <- bh_nutrients %>% mutate(DIN.mgL=(NOx+NH4)*14.007/1000,
                                        DIP.mgL=PO4*31/1000,
                                        TN.mgL=(TDN+PN)*14.007/1000,
                                        TP.mgL=(TDP+PP)*31/1000
                                        )

bh_nutrients$DIN.Flag <- ""
bh_nutrients$DIN.Flag[bh_nutrients$NOx.Flag=="<" | bh_nutrients$NH4.Flag=="<"] <- "<"
```

# Evaluate frequency of non-detects in the data

```{r}

# Evaluate frequency of non-detects
tmp <- bh_nutrients %>% 
  filter(between(month(TimeStamp),6,9)) %>% 
  select(NOx.Flag,NH4.Flag,DIN.Flag,PO4.Flag,TDN.Flag,
         PN.Flag,PO4.Flag,TDP.Flag,PP.Flag) %>% 
  pivot_longer(cols=everything(),names_to="parameter",values_to="value")
tmp$value[tmp$value==""] <- ">"

tmp <- tmp %>% 
  group_by(parameter,value) %>% 
  summarize(n=n(),.groups="keep") %>% 
  pivot_wider(id_cols=c("parameter","value"),names_from="value",values_from="n",
              values_fill=0)

names(tmp) <- c("parameter","LTMDL","GTMDL")
tmp <- tmp %>% mutate(percentND = 100*LTMDL/(LTMDL+GTMDL))

# Graph it

bh_nutrients %>% 
  filter(between(month(TimeStamp),6,9)) %>% 
  ggplot(aes(x=TimeStamp,y=NOx,color=factor(NOx.Flag)))+
  geom_point()+
  scale_y_log10()

bh_nutrients %>% 
  filter(between(month(TimeStamp),6,9)) %>% 
  ggplot(aes(x=TimeStamp,y=NH4,color=factor(NH4.Flag)))+
  geom_point()+
  scale_y_log10()

bh_nutrients %>% 
  filter(between(month(TimeStamp),6,9)) %>% 
  ggplot(aes(x=TimeStamp,y=PO4,color=factor(PO4.Flag)))+
  geom_point()+
  scale_y_log10()

bh_nutrients %>% 
  filter(between(month(TimeStamp),6,9)) %>% 
  ggplot(aes(x=TimeStamp,y=DIN.mgL,color=factor(DIN.Flag)))+
  geom_point()+
  scale_y_log10()

```
# Read the Clarity Data
```{r}

# Read the Boston Harbor Water Clarity Data
bh_clarity <- read.csv("C:/Users/jhagy/OneDrive - Environmental Protection Agency (EPA)/jim/Kreakie Collaborations/Boston Harbor/bh_secchi.csv",skip=5) %>% 
  filter(Surface.or.Bottom=="S") %>% 
  mutate(TimeStamp=as.POSIXct(Date.time..EASTERN.STANDARD.TIME.,format="%m/%d/%Y %H:%M",tz="Etc/GMT-5"))

keepCols <- c(1,2,3,4,5,14,7,12)
bh_clarity <- bh_clarity[,keepCols]
names(bh_clarity)[c(7,8)] <- c("TotalDepth.m","Secchi.Depth.m")

```

# Graph the summer data

```{r}

# Join nutrients and secchi data, keeping only data needed for POLR
bh_polr <- inner_join(
  bh_clarity %>% dplyr::select(Station.ID,TimeStamp,Secchi.Depth.m),
  bh_nutrients %>% dplyr::select(Station.ID,TimeStamp,DIN.mgL,DIP.mgL,TN.mgL,TP.mgL,Chla),
  by=c("Station.ID","TimeStamp"))


tmp <- bh_polr %>% filter(Station.ID %in% c(138,140)) %>% 
  filter(between(month(TimeStamp),6,8)) %>% 
  mutate(year=year(TimeStamp),
         Station.ID=factor(Station.ID))

# Summer TN appears to be decreasing at both stations since the mid 1990s, more quickly at Neponset
ggplot(tmp,aes(x=year,y=TN.mgL,color=Station.ID))+
  geom_point()+
  geom_smooth(method="lm")+
  ggtitle("Summer Total Nitrogen in Boston Harbor") +
  labs(y="mg N/L") +
  scale_y_log10()

# Summer DIN appears to be decreasing at both stations since the mid 1990s 
ggplot(tmp,aes(x=year,y=DIN.mgL,color=Station.ID))+
  geom_point()+
  geom_smooth(method="lm")+
  ggtitle("Summer Dissolved Inorganic Nitrogen in Boston Harbor") +
  labs(y="mg N/L") +
  scale_y_log10()

# Summer Chla appears to have decreased around 2000 - including the variability of Chla, 
# and then slowly increased thereafter but wioth less variability.  
ggplot(tmp,aes(x=year,y=Chla,color=Station.ID))+
  geom_point()+
  geom_smooth(method="loess")+
  scale_y_log10()+
  ggtitle("Summer Chl-a in Boston Harbor") +
  labs(y="µg/L")

tmp %>% group_by(Station.ID,year) %>% 
  summarize(p90=quantile(Chla,probs=0.9,na.rm=TRUE)) %>% 
  ggplot(aes(x=year,y=p90))+
  facet_grid(Station.ID~.)+
  geom_point()+
  geom_smooth(method="loess")+
  labs(x="Year",y="90th Percentile Chla µg/L")+
  ggtitle("Summer Chlorophyll-a Blooms in Boston Harbor Area")

summary(tmp)

```

# Select data to process for TSI calculation

Processing station 140, near Neponset River mouth (Dorchester) and station 138 (Inner Harbor/New England Aquarium)
https://www.mwra.com/harbor/graphic/harboronly_sampling_locations.png

```{r}

nut.variables <- c("NH4","NOx","TDN","PN","TDP","PP","PO4","Chla")
nut.flags <- c("NH4.Flag","NOx.Flag","TDN.Flag","PN.Flag","TDP.Flag","PP.Flag","PO4.Flag")

bh_polr <- inner_join(
  bh_clarity %>% dplyr::select(Station.ID,TimeStamp,Secchi.Depth.m),
  bh_nutrients %>% 
    dplyr::select(Station.ID,TimeStamp,all_of(nut.variables),all_of(nut.flags)),
  by=c("Station.ID","TimeStamp"))

# Establish upper and lower bounds for variables
bh_polr$NH4.upper <- bh_polr$NH4
bh_polr$NH4.lower[bh_polr$NH4.Flag=="<"] <- 0
bh_polr$NH4.lower[bh_polr$NH4.Flag==""] <- bh_polr$NH4[bh_polr$NH4.Flag==""]

bh_polr$NOx.upper <- bh_polr$NOx
bh_polr$NOx.lower[bh_polr$NOx.Flag=="<"] <- 0
bh_polr$NOx.lower[bh_polr$NOx.Flag==""] <- bh_polr$NOx[bh_polr$NOx.Flag==""]

bh_polr$TDN.upper <- bh_polr$TDN
bh_polr$TDN.lower[bh_polr$TDN.Flag=="<"] <- 0
bh_polr$TDN.lower[bh_polr$TDN.Flag==""] <- bh_polr$TDN[bh_polr$TDN.Flag==""]

bh_polr$PN.upper <- bh_polr$PN
bh_polr$PN.lower[bh_polr$PN.Flag=="<"] <- 0
bh_polr$PN.lower[bh_polr$PN.Flag==""] <- bh_polr$PN[bh_polr$PN.Flag==""]

bh_polr$PO4.upper <- bh_polr$PO4
bh_polr$PO4.lower[bh_polr$PO4.Flag=="<"] <- 0
bh_polr$PO4.lower[bh_polr$PO4.Flag==""] <- bh_polr$PO4[bh_polr$PO4.Flag==""]

bh_polr$TDP.upper <- bh_polr$TDP
bh_polr$TDP.lower[bh_polr$TDP.Flag=="<"] <- 0
bh_polr$TDP.lower[bh_polr$TDP.Flag==""] <- bh_polr$TDP[bh_polr$TDP.Flag==""]

bh_polr$PP.upper <- bh_polr$PP
bh_polr$PP.lower[bh_polr$PP.Flag=="<"] <- 0
bh_polr$PP.lower[bh_polr$PP.Flag==""] <- bh_polr$PP[bh_polr$PP.Flag==""]

for (i in seq(1,nrow(bh_polr))) {
  bh_polr$NH4.sample[i] <- runif(1,min=bh_polr$NH4.lower[i],max=bh_polr$NH4.upper[i])
  bh_polr$NOx.sample[i] <- runif(1,min=bh_polr$NOx.lower[i],max=bh_polr$NOx.upper[i])
  bh_polr$TDN.sample[i] <- runif(1,min=bh_polr$TDN.lower[i],max=bh_polr$TDN.upper[i])
  bh_polr$PN.sample[i] <- runif(1,min=bh_polr$PN.lower[i],max=bh_polr$PN.upper[i])
  bh_polr$PO4.sample[i] <- runif(1,min=bh_polr$PO4.lower[i],max=bh_polr$PO4.upper[i])
  bh_polr$TDP.sample[i] <- runif(1,min=bh_polr$TDP.lower[i],max=bh_polr$TDP.upper[i])
  bh_polr$PP.sample[i] <- runif(1,min=bh_polr$PP.lower[i],max=bh_polr$PP.upper[i])
    }

# Calculate variables that are calculated from reported values
bh_polr <- bh_polr %>% mutate(
  DIN=(NH4.sample+NOx.sample)*14.007/1000,
  TN=(TDN.sample+PN.sample)*14.007/1000,
  TP=(TDP.sample+PP.sample)*31/1000,
  DIP=PO4.sample*31/1000,
  DIN.Flag="",
  DIP.Flag="",
  TN.Flag="",
  TP.Flag="")

# If either of the components of a sum are below reporting limit, flag the sum.

bh_polr$DIN.Flag[bh_polr$NH4.Flag=="<" | bh_polr$NOx.Flag=="<"] <- "<"
bh_polr$DIP.Flag[bh_polr$PO4.Flag=="<" ] <- "<"
bh_polr$TN.Flag[bh_polr$TDN.Flag=="<" | bh_polr$PN.Flag=="<"] <- "<"
bh_polr$TP.Flag[bh_polr$TDP.Flag=="<" | bh_polr$PP.Flag=="<"] <- "<"

# Retain only the POLR variables and associated flags
polr.variables <- c("TN","TP","DIN","DIP","Secchi.Depth.m","Chla")
polr.flags <- c("TN.Flag","TP.Flag","DIN.Flag","DIP.Flag")
bh_polr <- bh_polr %>% 
  select(Station.ID,TimeStamp,all_of(polr.variables),all_of(polr.flags))

save(bh_polr,file="Boston_POLR_Data.Rdata")

```

# Graph the POLR variables

```{r}

tmp <- bh_polr %>% filter(Station.ID %in% c(138,140)) %>% 
  filter(between(month(TimeStamp),6,9)) 

tmp1 <- tmp %>% 
  select(Station.ID,TimeStamp,TN,TP,DIN,DIP) %>% 
  pivot_longer(
    cols = c("TN","TP","DIN","DIP"),
    names_to = "Param"
  )

tmp2 <- tmp %>% 
  select(Station.ID,TimeStamp,TN.Flag,TP.Flag,DIN.Flag,DIP.Flag) %>% 
  pivot_longer(
    cols = c("TN.Flag","TP.Flag","DIN.Flag","DIP.Flag"),
    names_to = "Param"
  ) %>% rename(Flag=value)
tmp3 <- tmp %>% 
  select(Station.ID,TimeStamp,Chla,Secchi.Depth.m) %>% 
  pivot_longer(
    cols = c("Secchi.Depth.m","Chla"),
    names_to = "Param"
  ) %>% 
  mutate(Flag="")

tmp <- cbind(tmp1,Flag=tmp2$Flag) %>% 
  rbind(tmp3) %>% 
  mutate(Flag=factor(Flag,levels=c("","<"),labels=c("none","< reporting limit")),
         Param=factor(Param,
                      levels=c("Chla","Secchi.Depth.m","TN","TP","DIN","DIP"),
                      labels=c("Chla (µg/L)","Secchi Depth (m)","TN (mg-N/L)",
                      "TP (mg-P/L)","DIN (mg-N/L)","DIP (mg-P/L)")))

rm(list=c("tmp1","tmp2","tmp3"))

plt <- ggplot(tmp,aes(x=TimeStamp,y=value,color=Flag))+
  facet_wrap(~Param,scales="free_y",ncol=2)+
  geom_point()+
  scale_y_log10()+
  labs(x="",y="")+
  theme(legend.position="top",legend.title=element_blank(),
        panel.grid=element_blank())

pdf(file="Boston_Harbor_TS.pdf",useDingbats = FALSE,paper="letter",width=7,height=9)
  plt
dev.off()
png(file="Boston_Harbor_TS.png",width=6,height=6,units="in",res=150)
  plt
dev.off()

tmp <- bh_polr %>% filter(Station.ID %in% c(138,140)) %>% 
  filter(between(month(TimeStamp),6,8)) %>% 
  mutate(year=year(TimeStamp),
         Station.ID=factor(Station.ID))

rm(tmp)

```

## Calculate TSI and trophic state probability

```{r}

load("Bayesian_POLR_2015update.(5-10-20).Rdata")
load("Bayesian_POLR_2010.(5-10-20).Rdata")
load("centeringParameters.Rdata")
load("Boston_POLR_Data.Rdata")

bh <- bh_polr %>% 
#  filter(Station.ID %in% c(138,140)) %>%
  filter(between(month(TimeStamp),6,9)) %>% 
  select(Station.ID,TimeStamp,TN,TP,DIN,DIP,Secchi.Depth.m,Chla) %>% 
  mutate(levelIII="NORTHEASTERN COASTAL ZONE") %>% 
  mutate(levelIII=factor(levelIII,levels=levelIIIregions))  # region variable must be a factor

bh <- bh[complete.cases(bh),]

# rename variables to match expected names
names(bh) <- c("Station.ID","TimeStamp","TN..mgN.L.","TP..mgP.L.","DIN..mgN.L.","DIP..mgP.L.","SECCHI_MEAN..m.","CHLA..ug.L.","levelIII")

Breaks_Chla_Q <- c(0,5,10,20,Inf)

# Define ordinal Chl-a class variable using quantile-based breaks
bh <- mutate(bh,TS_Chla_Q=cut(CHLA..ug.L., 
                                        breaks=Breaks_Chla_Q, 
                                        labels=c("Oligo", "Meso", "Eu", "Hyper")))

# Boston Harbor stations have chl-a that is oligo- to meso-trophic most of the time.
bh %>% group_by(TS_Chla_Q) %>% summarize(n=n()) %>% .[,2] / nrow(bh)


Alpha <- build_coefficient_matrix(Coeff.Summary.update.L3Eco)
regionMatrix <- build_region_matrix(bh,"levelIII")
centerParameters(bh,centeringParameters)
X <- cbind(SDD.C, TN.C, TP.C, DIN.C, DIP.C, regionMatrix)
TSI <- X %*% Alpha[,"mean"]
bh$TSI <- TSI

# Graph frequency histograms of 
bh$Station.Group[bh$Station.ID==138] <- "Inner Harbor"
bh$Station.Group[bh$Station.ID!=138] <- "Other Locations"

bh %>% mutate(era = cut(year(TimeStamp),breaks=c(1990,2000,2005,2010,2015,2021),
                        labels=c("1995-2000","2000-2005","2005-2010","2010-2015","2015-2020"))
              ) %>% 
  ggplot(aes(x=TSI))+
  geom_histogram(binwidth = 50,fill="lightblue")+
  facet_grid(era ~ .,scales="free_y")+
  scale_x_continuous(limits=c(-300,250),breaks=seq(-250,250,by=100))+
  labs(x="Trophic State Index",y="Frequency")+
  geom_vline(xintercept=Coeff.Summary.update.L3Eco[1:3,"mean"],linetype="dotted")+
  theme_classic()+
  theme(legend.position="top",strip.background = element_blank())

ggplot(bh,aes(x=year(TimeStamp),y=TSI,color=TS_Chla_Q))+
  geom_point()+
  scale_y_continuous(limits=c(-300,250),breaks=seq(-300,250,by=100))+
  scale_x_continuous(breaks=c(1995,2000,2005,2010,2015,2020))+
  labs(y="Trophic State Index",x="")+
  geom_hline(yintercept=Coeff.Summary.update.L3Eco[1:3,"mean"],linetype=2)+
  theme_classic()


tmp <- calculate_ts_probabilities(Coeff.Summary.update.L3Eco,bh$TSI)

bh <- cbind(bh,tmp)
save(bh,file="BostonHarborTSI.Rdata")
write.csv(bh,file="BH_TSI.csv",row.names=FALSE)
```

# Evaluate agreement between TSI and Chla

```{r}

load("BostonHarborTSI.Rdata")

regionEval <- build_region_matrix(bh,"levelIII")
Pred.CatAll <- findPredictedClass.TSI(bh,Coeff.Summary.update.L3Eco)

CM.TS.Multilevel <- confusionMatrix(Pred.CatAll,bh$TS_Chla_Q)
CM.TS.Multilevel


# Classify Observations
cut_pts <- Coeff.Summary.update.L3Eco[1:3,"mean"]
bh <- bh %>% mutate(TS_Predicted=cut(TSI,breaks=c(-Inf,cut_pts,Inf),labels=levels(bh$TS_Chla_Q)))

bh %>% filter(TS_Chla_Q=="Oligo" & TS_Chla_Q != TS_Predicted) %>% .[,13:16] %>% summary()

tmp <- bh %>% 
  select(year,pMesotrophic,pOligotrophic,pEutrophic,pHypereutrophic) %>% 
  pivot_longer(cols=starts_with("p"),names_to="ProbName",values_to="Prob") %>% 
  mutate(ProbName=factor(ProbName,levels=c("pOligotrophic","pMesotrophic","pEutrophic","pHypereutrophic"),
         labels=c("Oligotrophic","Mesotrophic","Eutrophic","Hypereutrophic")))


ggplot(tmp,aes(x=year,y=Prob))+
  facet_wrap(ProbName ~ .,ncol=2)+
  geom_point()+
  geom_smooth()+
  scale_x_continuous(breaks=seq(1995,2015,by=5))
  
```

# Examine How TSI changed in Boston Harbor and compare with how Chl-a changed 

Change in TSI reflects "macroscope" prediction of hos chl-a would be expected to change, whereas change in Chlorophyll-a reflects observed change.  

```{r}
ggplot(bh,aes(x=year,y=TSI))+
  geom_point()+
  geom_smooth()

install.packages("quantreg")
library("quantreg")

# Model year trend in median TSI
fit1 <- rq(TSI ~ year, tau = .5, data = bh)

# Model year trend in median TSI
fit2 <- rq(TSI ~ year, tau = .9, data = bh)
fit2

ggplot(bh,aes(x=year,y=TSI))+
  geom_point()+
  geom_abline(aes(slope=fit2$coefficients[2],intercept=fit2$coefficients[1]))

ggplot(bh,aes(x=year,y=CHLA..ug.L.))+
  geom_point()+
  scale_y_log10()

```

Fit a QGAM through summer Chl-a data in Boston Harbor

```{r}

fit90 <- qgam(CHLA..ug.L. ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.9)
fit50 <- qgam(CHLA..ug.L. ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.5)
fit10 <- qgam(CHLA..ug.L. ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.1)

xSeq <- data.frame(year=seq(1995,2020))
pred90 <- predict(fit90,newdata=xSeq,se=TRUE)
pred50 <- predict(fit50,newdata=xSeq,se=TRUE)
pred10 <- predict(fit10,newdata=xSeq,se=TRUE)

tmp <- data.frame(year=xSeq$year,
                  q90=pred90$fit,
                  se90=pred90$se.fit,
                  q50=pred50$fit,
                  se50=pred50$se.fit,
                  q10=pred10$fit,
                  se10=pred10$se.fit) %>% 
                  mutate(year=as.POSIXct(paste0(as.character(year),"-07-01"),format="%Y-%m-%d"))


pal <- brewer.pal(4,"Set2")
chla_Qgam_plt <- ggplot()+
     geom_point(data=bh,aes(x=TimeStamp,y=CHLA..ug.L.),shape=1,size=2,color="grey")+
     geom_ribbon(data=tmp,aes(x=year,ymin=q90-se90,ymax=q90+se90),fill=pal[1])+
     geom_line(data=tmp,aes(x=year,y=q90))+
     geom_ribbon(data=tmp,aes(x=year,ymin=q50-se50,ymax=q50+se50),fill=pal[2])+
     geom_line(data=tmp,aes(x=year,y=q50))+
     geom_ribbon(data=tmp,aes(x=year,ymin=q10-se10,ymax=q10+se10),fill=pal[3])+
     geom_line(data=tmp,aes(x=year,y=q10))+
     #theme_classic()+
     scale_x_datetime(breaks="5 years",date_labels="%Y")+
     scale_y_log10(breaks=c(3,5,10,20,30))+
     labs(x="",y="µg L-1")
chla_Qgam_plt  

pdf(file="QGAM figure.pdf",width=4,height=4,useDingbats = FALSE,paper="letter")
  chla_Qgam_plt  
dev.off()
```

Fit a QGAM through TSI data in Boston Harbor

```{r}
bh$year <- year(bh$TimeStamp)

fit90 <- qgam(TSI ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.9)
fit50 <- qgam(TSI ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.5)
fit10 <- qgam(TSI ~ s(year,k=10,bs="ad"),
            data=bh,
            qu=0.1)

xSeq <- data.frame(year=seq(1995,2020))
pred90 <- predict(fit90,newdata=xSeq,se=TRUE)
pred50 <- predict(fit50,newdata=xSeq,se=TRUE)
pred10 <- predict(fit10,newdata=xSeq,se=TRUE)

tmp <- data.frame(year=xSeq$year,
                  q90=pred90$fit,
                  se90=pred90$se.fit,
                  q50=pred50$fit,
                  se50=pred50$se.fit,
                  q10=pred10$fit,
                  se10=pred10$se.fit) %>% 
                  mutate(year=as.POSIXct(paste0(as.character(year),"-07-01"),format="%Y-%m-%d"))


pal <- brewer.pal(4,"Set2")
TSI_Qgam_plt <- ggplot()+
     geom_point(data=bh,aes(x=TimeStamp,y=TSI),shape=1,size=2,color="grey")+
     geom_ribbon(data=tmp,aes(x=year,ymin=q90-se90,ymax=q90+se90),fill=pal[1])+
     geom_line(data=tmp,aes(x=year,y=q90))+
     geom_ribbon(data=tmp,aes(x=year,ymin=q50-se50,ymax=q50+se50),fill=pal[2])+
     geom_line(data=tmp,aes(x=year,y=q50))+
     geom_ribbon(data=tmp,aes(x=year,ymin=q10-se10,ymax=q10+se10),fill=pal[3])+
     geom_line(data=tmp,aes(x=year,y=q10))+
     #theme_classic()+
     scale_x_datetime(breaks="5 years",date_labels="%Y")+
     scale_y_continuous(limits=c(-300,200),breaks=seq(-300,200,by=50))+
     labs(x="",y="Trophic State Index")


pdf(file="QGAM TSI figure.pdf",width=4,height=4,useDingbats = FALSE,paper="letter")
  TSI_Qgam_plt
dev.off()
```

# Q-Q plots
```{r}
load("coastalTSI.Rdata")
load("BostonHarborTSI.Rdata")
load("Bayesian_POLR_2015update.(5-10-20).Rdata")

library("wesanderson")
probs <- seq(0.05,0.95,by=0.05)

q.NCCA.TSI <- quantile(coastalTSI$TSI,probs=probs)
q.NE.TSI <- quantile(coastalTSI$TSI[coastalTSI$levelIII=="NORTHEASTERN COASTAL ZONE"],probs=probs)
q.BH.TSI <- quantile(bh$TSI,probs=probs)
pal <- wes_palette("Zissou1", 10, type = "continuous")

data.frame(q.NCCA.TSI,q.BH.TSI,q.NE.TSI,Quantile=probs) %>% 
  ggplot() +
  geom_point(aes(x=q.NE.TSI,y=q.BH.TSI,color=Quantile),shape=17)+
  geom_point(aes(x=q.NCCA.TSI,y=q.BH.TSI,color=Quantile)) +
  scale_color_gradientn(colours = pal)+
  scale_x_continuous(limits=c(-400,400))+
  scale_y_continuous(limits=c(-400,400))+
  geom_abline(scope=1,intercept=0, linetype=2)+
  geom_vline(aes(xintercept=Coeff.Summary.update.L3Eco[1,"mean"]),color="grey")+
  geom_vline(aes(xintercept=Coeff.Summary.update.L3Eco[2,"mean"]),color="grey")+
  geom_vline(aes(xintercept=Coeff.Summary.update.L3Eco[3,"mean"]),color="grey")+
  geom_hline(aes(yintercept=Coeff.Summary.update.L3Eco[1,"mean"]),color="grey")+
  geom_hline(aes(yintercept=Coeff.Summary.update.L3Eco[2,"mean"]),color="grey")+
  geom_hline(aes(yintercept=Coeff.Summary.update.L3Eco[3,"mean"]),color="grey")+
  theme_classic()+
  labs(y="Boston Harbor TSI",x="NCCA or Northeast TSI")
```

# Q-Q Plot of TSI and Chlorophyll-a by period

```{r}
yrs <- c(1995,2000)
q.BH.TSI <- quantile(bh$TSI[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
q.BH.Chla <- quantile(bh$CHLA..ug.L.[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
interval <- "1995-2000"
tmp <- data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval)

yrs <- c(2000,2005)
q.BH.TSI <- quantile(bh$TSI[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
q.BH.Chla <- quantile(bh$CHLA..ug.L.[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
interval <- "2000-2005"
tmp <- rbind(tmp,data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval))

yrs <- c(2005,2010)
q.BH.TSI <- quantile(bh$TSI[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
q.BH.Chla <- quantile(bh$CHLA..ug.L.[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
interval <- "2005-2010"
tmp <- rbind(tmp,data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval))

yrs <- c(2010,2015)
q.BH.TSI <- quantile(bh$TSI[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
q.BH.Chla <- quantile(bh$CHLA..ug.L.[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
interval <- "2010-2015"
tmp <- rbind(tmp,data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval))

yrs <- c(2015,2019)
q.BH.TSI <- quantile(bh$TSI[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
q.BH.Chla <- quantile(bh$CHLA..ug.L.[between(year(bh$TimeStamp),yrs[1],yrs[2])],probs=probs)
interval <- "2015-2019"
tmp <- rbind(tmp,data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval))

q.BH.TSI <- quantile(coastalTSI$TSI,probs=probs)
q.BH.Chla <- quantile(coastalTSI$CHLA..ug.L.,probs=probs)
interval <- "NCCA"
tmp <- rbind(tmp,data.frame(q.BH.TSI,q.BH.Chla,Quantile=probs,interval=interval))

pal <- brewer.pal(6,"Set2")

ggplot(tmp,aes(x=q.BH.TSI,y=q.BH.Chla,color=interval))+
  geom_point()+
  geom_smooth(se=FALSE)+
  #scale_y_log10(breaks=c(2.5,5,10,20))+
  scale_color_manual(values=pal)+
  #scale_x_continuous(limits=c(-250,50))+
  geom_vline(aes(xintercept=Coeff.Summary.update.L3Eco[1,"mean"]),color="grey")+
  #geom_vline(aes(xintercept=Coeff.Summary.update.L3Eco[2,"mean"]),color="grey")+
  #geom_hline(aes(yintercept=5),color="grey")+
  #geom_hline(aes(yintercept=10),color="grey")+
  #geom_hline(aes(yintercept=20),color="grey")+
  theme_classic()+
  labs(x="Trophic State Index",y="Chlorophyll-a (µg/L",color="")




```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
