---
title: "Demonstrate how fill-in imputation using ROS works"
author: "J. Hagy"
date: "7/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Demonstrate fill-in methods using ROS in NADA

ros is an implementation of a Regression on Order Statistics (ROS) designed 
for multiply censored analytical chemistry data.
The method assumes data contains zero to many left-censored 
less-than) values.

ros is part of the NADA package

Lopaka Lee (2020). NADA: Nondetects and Data Analysis for Environmental Data. R package version 1.6-1.1.
https://CRAN.R-project.org/package=NADA

```{r}

# Use the NADA package ... developed with Denis Helsel
library(NADA)
library(tidyverse)

```

# Create a log-normally distributed random variable

You can also embed plots, for example:

```{r}
# Create log-normally distributed data data set and proposed censoring limit
set.seed(5420)
myRndData <- exp(sort(rnorm(1000)))

# define values less than 0.2 as being "below detection limit"
censored <- myRndData<0.2
summary(censored)
# about 6.1% of the data are censored.

# put the data in a data frame and graph is, showing the data
# that I'm going to define as <MDL
tmp <- data.frame(x=myRndData,censored=censored)

ggplot(tmp,aes(x=x,fill=censored))+
  geom_histogram(binwidth=0.1,boundary=1)+
  scale_x_log10()+
  ggtitle("Distribution Before Censoring")

```

# Censor the data 
Now actually replace data below with MDL with the MDL and plot showing 
the left-censored distribution

```{r}
# Replace the data below 0.2 with 0.2  
tmp$x_cens <- pmax(tmp$x,0.2)

# now there are more numbers in the lowest bin, because they are all the 
# censored values
ggplot(tmp,aes(x=x_cens,fill=censored))+
  geom_histogram(binwidth=0.1,boundary=1)+
  scale_x_log10()+
  ggtitle("Distribution After Censoring")

```
# Use ROS 
ROS will fill-in the censored tail of the distribution 
with imputed values.
```{r}

# use regression on order statistics (ROS) to fill-in
# the distribution of values in the left tail to replace
# the values that are censored.  
myros = ros(tmp$x_cens, tmp$censored) 
myros.df <- as.data.frame(myros)

ggplot(myros.df,aes(x=modeled,fill=censored))+
  geom_histogram(binwidth=0.1,boundary=1)+
  scale_x_log10()+
  ggtitle("Distribution with Imputed values in the left tail")

```

